---
title: "SimFinR Review"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a reviw of the package [SimFinR](https://github.com/msperlin/simfinR/).

This package was released to [CRAN](https://cran.r-project.org/web/packages/simfinR/index.html) at the end of 2019 for v 0.1.0, and is a wrapper for the [SimFin](https://simfin.com/) project, whose mission statement of of prodiving cleaned, standardized data from EDGAR and stock information to help reduce barriers for learning algorithms and new investors. If you are interested in the SimFin API information, that can be found [here](https://simfin.com/api/v1/documentation/).

If you would like this RMarkdown document, that can be found [here](https://github.com/willcrouch/SimFinReview).

Please download the .rmd file and play around with this package, it looks like it will be a solid addition to any finance R user.

```{r echo = FALSE}
library(simfinR)
```

The first step you will need to use the package is to acquire an [API](https://simfin.com/data/api) key.  All you will need to do is register an account and then going to the API tab for your account's API key.  The API allows for 2,000 calls per day for a free user, and signing up for SimFin+ for more calls.  

We will use an API from a dummy account.

```{r}
APIkey = "zgPAQaE9G5MdYHDFTbeKtQMR3u4iUgp5"
```

First we are going to get a list of all the companies available on SimFin.

```{r}
CompList = simfinR_get_available_companies(APIkey)

# see top companies
head(CompList)
```

## Company Information

The data consist of 3 fields, a unique serial key for each company, the ticker symbol, and the company name.  For this review, we will look at Pepsi.  And if you want to see a good [summary of their data](https://simfin.com/data/companies/218222).

```{r}
CompList[CompList$name == "PEPSICO INC",]

CompanyID = 218222
TickerSymbol = "PEP"
CompanyName = "PEPSICO INC"
```

SimFin provides the information for the company and name if you just have the serial key.  Not sure if the specific use case for that call, but hey, it's available!

```{r}
simfinR_get_info_company(CompanyID, APIkey)
```

```{r}
simfinR_id_to_name(CompanyID, APIkey)
```

## Financial Statements

We can call the fill document list for all available statemets that SimFin has available on Pepsi.

```{r}
DocumentList = simfinR_get_available_statements(CompanyID, APIkey, silent = TRUE)

# Look at a couple of the entries in the list
head(DocumentList)
```

We can see that SimFin provides not just EDGAR filings, but also calculated fields like trailing twelve months [TTM] information and quarterly information depending on how its presented from EDGAR.  With Pepsi their 2nd 10Q is presented in 3 months & 8 months for Income Statement but just 8 months for Cash Flow.

The data returned is the period of the reported data, fiscal year, calculated flag, statement type, serial key for the comapny, and company name.

If we want to pull out just the financial statements, we can do that and call out the specific types.
pl = income statement (profit/loss)
bs = balance sheet
cf = cash flow

```{r}
FinStatements = simfinR_get_fin_statements(
  CompanyID,
  APIkey,
  type_statements = c("pl", "bs", "cf"),
  periods = "FY",
  years = 2018,
  cache_folder = "simfin_cache"
)

head(FinStatements)
```

This will pull out all the in the financial statements:
* Company Name
* Company Sector
* Statement Type
* Fiscal Period
* Year
* Reference Date
* Account Name 
* Account Value

JSON parase data from the call:
* tid
* uid
* Display Level
* Check Possible

You will see that this does not return Balance Sheet data for Pepsi.  This is because they balance sheet data is stored by quarter, and there is no FY level return for that information.  So we would get that with this call:
```{r}
PEPBalSheet = simfinR_get_fin_statements(
  CompanyID,
  APIkey,
  type_statements = "bs",
  periods = c("Q1","Q2","Q3","Q4"),
  years = 2018,
  cache_folder = "simfin_cache"
)

head(PEPBalSheet)
```


We can also pull a specific financial statement and its information.

```{r}
SingleStatement = simfinR_get_single_fin_statement(
  CompanyID,
  type_statement = "pl",
  period_in = "FY",
  year = 2018,
  APIkey,
  cache_folder = "simfim_cache" # This is just a named cache, nothing preset
  )

head(SingleStatement)
```

## Stock History

You can call the stock price history for a company, or more than one company if you want to pull a list.
```{r}
StockPrice = simfinR_get_price_data(CompanyID, APIkey)

head(StockPrice)
```
This call provides standard stock information:
* Date
* Adjusted Close
* Split Coeffienct
* Share count in class
* Class Name
* Call Type
* Currency
* Company Name


You can also call just a specific company's stock information.  This method makes use of a cache folder for easier handling.
```{r}
SingleStock = simfinR_get_single_price_data(CompanyID, APIkey, cache_folder = "simfim_cache")

head(SingleStock)
```
This call provides the same stock information:
* Date
* Adjusted Close
* Split Coeffienct
* Share count in class
* Class Name
* Call Type
* Currency
* Company Name

Let's take a look at the stock call with Coke as well.
```{r}
# Coke ID = 156024
StockList = c(218222,156024)

StockPricePC = simfinR_get_price_data(StockList, APIkey)

head(StockPricePC)
```
This call provides the same stock information, but both in the same return.
```{r}
nrow(StockPrice)
nrow(StockPricePC)
```

## Other call

You can also use the API to help you out with some time periods if you are not used to financial timeframes.  This call provides the last date for a given period and year
```{r}
Time_Period = "Q1"
Time_Year = 2018
simfinR_period_to_date(period_in = Time_Period, year_in = Time_Year)
```

